Flaring star-planet interactions in star-planet systems in Kepler and TESS

### 2023-05-10 src/tex/ms.tex
[showyourwork] overleaf sync

Tidal interaction can also lead to enhanced activity, globally, and locally. Locally, we expect to see periodicity with half the orbital period of the planet, corresponding to the two tidal bulges forming in the stellar envelope.

@@

We applied the same technique as described in Section~\ref{sec:methods:adtest} to determine the significance of a local tidal SPI signal. Following~\citet{ilic2022tidal}, we compare the resulting $p$-values with three different models of tidal interaction in Fig.~\ref{fig:tidal} and list the derived values in Table~(XXXX). The first model measures the torque $\partial L / \partial t$, exerted by angular momentum transfer from star to planet or from planet to star~(\citealt{penev2012constraining}, their Eqns. 1 and 2). We choose the tidal quality factor $Q_*=10^7$, but our results do not depend on the choice of $Q_*$. We also make the simplifying assumption that the rotation period of the convective envelope is the same as the surface rotation period measured as described in Section~\ref{sec:data:rotationperiods}, ignoring differential rotation effects. \textcolor{red}{Why doesn't the torque depend on the stellar rotation velocity, except for the sign of interaction?}. The second model considers the tidal dissipation timescale -- the shorter the timescale, the higher the power of the interaction~\citep{zahn1977tidal}. Following~\citet{ilic2022tidal}, we adopt Eqn.2 from~\citep{albrecht2012obliquities} for the tidal dissipation timescale in the convective envelope. We assume that all stars in our sample have a convective envelope because they flare, and flaring requires a dynamo-generated magnetic field, which can only operate in stars with convective envelopes. Most of our stars are clearly below the Kraft break at spectral type F5~(XXXX), above which stars have a radiative envelope, except for KOI-12, which is an F5 sub-giant~\citep{frasca2016activity}. The third model uses the gravitational perturbation as a proxy for the interaction~\citep{cuntz2000stellar}. All models require the planetary and stellar mass to be known, which we collect from the literature. For missing planetary masses, radii are known, so we can use the mass-radius relations in~\citet{chen2017probabilistic}. We list the masses in Table~(XXXX).

@@

Among the systems, HIP 67522 appears with tentative signs of both tidal and magnetic interaction (measured and expected in the different scenarios). On the one hand, since the flares in HIP 67522 mostly appear close to the transit phase, and not during eclipse, this may favor magnetic interaction, and the tidal signal would only be an artifact. On the other hand, the narrow range in which most of the flares on HIP 67522 occurred favor a well-located tidal bulge. It is conceivable that both effects play a role at the same time.

>>>>

\begin{table*}
\footnotesize
\movetableright=-20mm
\script{paper_table_tidal.py}
\caption{Flaring tidal star-planet interaction.}
\input{output/table_tidal.tex}
    \label{tab:maintable_der}
 \tablerefs{\input{output/table_tidal_bibstring.tex}}
\end{table*}

Tidal interaction can also lead to enhanced activity, globally, and locally. Locally, we expect to see periodicity with half the orbital period of the planet, corresponding to the two tidal bulges forming in the stellar envelope. XXXX tidal theory here XXXX

@@

We applied the same technique as described in Section~\ref{sec:methods:adtest} to determine the significance of a local tidal SPI signal. Following~\citet{ilic2022tidal}, we compare the resulting $p$-values with three different models of tidal interaction as proxies for expected excess flaring due to tides in Fig.~\ref{fig:tidal}. We list the derived values in Table~(XXXX).

The first model measures the torque $\partial L / \partial t$, exerted by angular momentum transfer from star to planet or from planet to star~(\citealt{penev2012constraining}, their Eqns. 1 and 2). We choose the tidal quality factor $Q_*=10^7$. Our results do not depend on the choice of $Q_*$, as we only need to know the tidal torque down to a proportionality factor. We also make the simplifying assumption that the rotation period of the convective envelope is the same as the surface rotation period measured as described in Section~\ref{sec:data:rotationperiods}, ignoring differential rotation effects. \textcolor{red}{Why doesn't the torque depend on the stellar rotation velocity, except for the sign of interaction?}.

The second model considers the tidal dissipation timescale -- the shorter the timescale, the higher the power of the interaction~\citep{zahn1977tidal}. Following~\citet{ilic2022tidal}, we adopt Eqn.2 from~\citep{albrecht2012obliquities} for the tidal dissipation timescale in the convective envelope. We assume that all stars in our sample have a convective envelope because they flare, and flaring requires a dynamo-generated magnetic field, which can only operate in stars with convective envelopes. Most of our stars are clearly below the Kraft break at spectral type F5~(XXXX), above which stars have a radiative envelope, except for KOI-12, which is an F5 sub-giant~\citep{frasca2016activity}.

The third model uses the gravitational perturbation as a proxy for the interaction~\citep{cuntz2000stellar}. All models require the planetary and stellar mass to be known which we collect from the literature. For missing planetary masses, radii are known, so we can use the mass-radius relations implemented in \texttt{forecaster}~\citep{chen2017probabilistic}. GJ 3323, GJ 674 and GJ 3082 were also missing stellar mass estimates, which we obtained from the \citep{mann2015how, mann2016erratum} relations between absolute Ks magnitude~\citep{skrutskie2006two} and stellar mass for M dwarfs using the distances from \citet{bailer-jones2021estimating}. We list the masses in Table~(XXXX).

@@

Among the systems, HIP 67522 appears with tentative signs of both tidal and magnetic interaction (measured and expected in the different scenarios). On the one hand, since the flares in HIP 67522 mostly appear close to the transit phase, and not during eclipse, this may favor magnetic interaction, and the tidal signal would only be an artifact. On the other hand, the narrow range in which most of the flares on HIP 67522 occurred might favor a well-located tidal bulge. It is conceivable that both effects play a role at the same time.
